Stability of variational quantum dynamics with Gaussian approximants

### 2013-08-06 paper.tex
cancellation of weights explained

%% Phase space expansion as a linear operator

@@

%% SVE lemma

% Hypothesis: The singular values of $|Γ〉$ are 0 and 1, and the space belonging to 1 comprises the entire functions.
% 
% There are two ways one could set about proving the Hypothesis.  Firstly, one could show that any function orthogonal to the Gaussian analytics annihilates $G$.  Otherwise, one could show that the norms of all Gaussian analytics are scaled by the same factor going to F
% ock space, and that they are a pre-image of Fock space.  The second approach sounds simpler.
% 
% This won't be a rigourous argument: it could turn out that there's another class of expansions with half the norm, and the entire functions have components belonging to the 0 singular value.  My numerical experiment suggests otherwise; let's ignore that possibility for now.
% 
% Let $f$ be a normalised entire function, in Glauber's sense.  I.e., 
% $$f(z)=∑c_n{zⁿ\over (n!)^½},$$
% for some normalised vector $c$.
% 
%   Our expansion function is 
% $$g(α)=f(α)e^{-½|α|²}.$$
% If $f$ has the Taylor expansion
% $$f(z)=∑_n c_n{zⁿ\over(n!)^½},$$
% then 
% $$Gg = ∑_n c_n|n〉.$$
% The functions $zⁿ$ for $n≥0$ span the entire functions (or their span is a dense subset, or something like that), so let's find the norm of 
% $$g(α)=e^{-½|α|²}{αⁿ\over(n!)^½},$$
% the Glauber preimage of $|n〉$.  We have 
% $$|g(α)|²={1\over n!}e^{-|α|²}|α|^{2n},$$
% so the squared norm is
% $$‖g‖²={1\over n!}\int ₀^∞2πr^{2n+1}e^{-r²}\,dr={2π\over n!}\int ₀^∞r^{2n+1}e^{-r²}\,dr.$$
% According to Wolfram,
% $$\int ₀^∞r^{2n+1}e^{-r²}\,dr={n!\over 2},$$
% so
% $$‖g‖=π^½.$$
% 
% Assuming that there is no way to expand a state with a smaller norm than an entire function, this does it.  Or at least I think so.  I should look up exactly what the uniqueness properties are for singular value expansions in infinite dimensional spaces.

%% Consequences of lemma

@@

%% Finite phase space

@@

%% Picard plots

@@

It seems that, when two amplitudes are very close, the solutions where the weights cancel out are better fits than the smooth ones—that seems to be the case in the L-curve experiments for the expansion problem.  To avoid the amplitudes clustering, I need to find out how good a fit is possible with evenly spaced ampltiudes, and set the regularisation parameter so that the residual is on that order.  In fact, I already know how large the phase space expansion should be: the function being sampled should have a norm around 1.

@@

%% Adapting the sample points to maintain condition

@@

%% Discrete phase space

@@

%% Why grids don't work, and variation necessary

@@

%% convergence to the initial state

@@

%% Shape of Hilbert space trajectories

@@

>>>>

\beginsection{The phase space expansion operator}

@@

%% Consequences of Γ being a projection

@@

\beginsection{Truncated phase space}

@@

\beginsection{Regularisation of the truncated expansion problem}
%% Picard plots

@@

To avoid the amplitudes clustering, I need to find out how good a fit is possible with evenly spaced ampltiudes, and set the regularisation parameter so that the residual is on that order.  In fact, I already know how large the phase space expansion should be: the function being sampled should have a norm around 1.

@@

A common observation when states are expanded over an irregular grid is that, when two sample amplitudes are close together, the weights tend to be large, and cancel out.  This is not at all surprising when one considers that the coherent states $|α〉$ and $|α+h〉$ become nearly parallel in Hilbert space as $h→0$.  The angle between these vectors satisfies
$$|〈α|α+h〉|=e^{-½|h|²}=‖|α〉‖‖|β〉‖\cos θ=\cos θ,$$
and equating Taylor series gives
$$θ=|h|+O(h²).$$
Locally, phase space has the same geometry as the surface of the unit sphere.  The figure demonstrates why weights of nearly parallel (or antiparallel) vectors tend to cancel out, and the second figure shows how this looks in phase space.

\topinsert \XeTeXpicfile geometry-1.pdf
\it A 2 dimensional section through Hilbert space, where fixed vector $|Ψ〉$ is expanded as $|Ψ〉=a|α〉+b|β〉$.  As $|β〉$ becomes more parallel with $|α〉$, the components grow and cancel.  This occurs for all vectors outside the shaded segments, which shrink to nothing when $|α〉$ becomes parallel to $|β〉$.
\endinsert

\topinsert \XeTeXpicfile cancel.pdf width \hsize
\it The vacuum state was expanded over two coherent states, with amplitudes $±h$.  The amplitudes and the entire expansion of the least-squares fit are plotted in the first row of images.  The norm of the expansion vector, and the residual of the expansion, are graphed above as black lines.  As the amplitudes converge, the residual decreases—the leftmost phase space plot is very nearly exact—while the norm of the expansion stays around 1.  This coincides with the amplitudes passing through bright areas of the images, where the entire expansion has support.

The second row of images, and the red lines on the plot, show similar expansions of the normalised state parallel to $|0.1〉-|-0.1〉$.  This is almost exactly parallel to the derivative of the vacuum state with respect to the real part of its amplitude.  Again, the leftmost image is nearly an exact expansion.  This time, the entire expansion does not have support around $α=0$.  As the expansion amplitudes converge, the residual decreases, but the norm of the expansion vector explodes.  As explained in the text, the distance between two small amplitudes in phase space is very nearly the angle between the ket vectors for the corresponding coherent states.  Thus this is exactly the expansion depicted in the last figure.

If the two sample amplitudes were part of an irregular expansion, and there happened to be no other sample amplitudes nearby, a small component along the direction of the second state would dominate the expansion coefficients at these points.  But, given those sample amplitudes, that would be the only way to represent such a component.
\endinsert

\beginsection{Adapting the sample amplitudes to preserve condition}

@@

\beginsection{Discrete phase space}

@@

\beginsection{Ways the problem has been simplified for this paper}

@@

\beginsection{Regularising the inital fit equation}

@@

\beginsection{Quantum dynamics in Hilbert space}

@@

\beginsection{Quantum dynamics in phase space}
